Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-10-22 draft.tex
towards computing the center of CE cochains

The notation $\int_{S^{n-1}}\UU_n\fg$ produces a factorization algebra on $\RR$.
Indeed,
\begin{align*}
\int_{S^{n-1}}\UU_n\fg 
&\simeq \cliels(\Omega^\bu_c(\RR_{> 0} \times S^{n-1}) \otimes \fg)\\
&\simeq \cliels(\Omega^\bu_c(\RR_{> 0}) \otimes H^\bu(S^{n-1}) \otimes \fg)
\end{align*}
because $(n-1)$-sphere is formal over the rationals (or reals). 
We thus see that 
\[
\int_{S^{n-1}}\UU_n\fg \simeq \UU_1 (H^\bu(S^{n-1}) \otimes \fg)
\]
The cohomology of the $(n-1)$-sphere is 
\[
R_n = k[\epsilon_{n-1}]/(\epsilon_{n-1}^2),
\] 
where $|\epsilon_{n-1}| = n-1$.
Since $\UU_1 (H^\bu(S^{n-1}) \otimes \fg)$ is manifestly equivalent to the dg algebra~$U(\fg \otimes R_n)$,
we find that 
\[
\int_{S^{n-1}}\UU_n\fg \simeq U(\fg \otimes R_n)
\]
as $E_1$ algebras.
Recall as well that $\UU_n \fg$ is quasi-isomorphic to $\Sym(\fg[1-n])$ as cochain complexes.

We have thus reduced the situation to familiar dg algebra via the equivalence
\[
{\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n\fg}}(\UU_n \fg, (\UU_n \fh)_f) 
\simeq 
{\rm Mor}_{\LMod_{U(\fg \otimes R_n)}} (\Sym(\fg[1-n]), (\Sym(\fh[1-n]))_f).
\]
By the inclusion of the unit $k \hookrightarrow R_n = k \oplus k[1-n]$,
we obtain a canonical inclusion $i: U\fg \to U(\fg \otimes R_n)$ of dg associative algebras,
and so
\[
k \otimes_{U\fg} U(\fg \otimes R_n) \simeq \Sym(\fg[1-n]).
\]
Base change along $i$ thus gives rise to an equivalence 
\[
{\rm Mor}_{\LMod_{U\fg}} (k, (\Sym(\fh[1-n]))_f) \simeq {\rm Mor}_{\LMod_{U(\fg \otimes R_n)}} (\Sym(\fg[1-n]), (\Sym(\fh[1-n]))_f)
\]
of underlying cochain complexes, i.e. an equivalence after applying the forgetful functor. 
The left hand side is modeled by $\clies(\fg,(\Sym(\fh[1-n]))_f)$, 
the Lie algebra cochains of $\fg$ with coefficients in~$(\UU_n \fh)_f$, as desired.
Hence the map $\theta_{\bf f}$ is a quasi-isomorphism on any disk, as needed.
\end{proof}

@@

When $n=1$ the quasi-isomorphism of the above proposition produces the well-known quasi-isomorphism
\[
{\rm Hoch}^\bu(U \fg , U \fg) \simeq \clie^\bu(\fg , \Sym(\fg)).
\]
See~\cite{LodayCyclicBook} for discussion of Hochschild (co)homology for such ``almost commutative'' algebras.

%As a corollary of the above proposition, we obtain a concrete description for $\EE_n$-deformations of $\UU_n \fg$ when $\fg$ is an ordinary (strict) Lie algebra. 
Consider the situation of an ordinary Lie algebra~$\fg$.
When $\fg$ is semi-simple, the enveloping $\EE_1$ algebra $\UU_1 \fg$ is rigid in the sense that it admits no nontrivial deformations. 
This claim follows from the above equivalence (\ref{eqn:def}) together with the fact that $H_{\rm Lie}^2(\fg , \Sym(\fg)) = 0$ for $\fg$ is semi-simple by Whitehead's theorem.

For higher $n$, the $\EE_n$-enveloping algebra is no longer rigid, even in the case of semi-simple Lie algebras. 
For instance, consider the case $n=2$.
\begin{cor}
Deformations of $\UU_2 \fg$ are in bijective correspondence with the group
\[
H^3_{\rm Lie}(\fg) \oplus H^2_{\rm Lie}(\fg , \fg) \oplus H^1_{\rm Lie}(\fg , \wedge^2 \fg) \oplus \wedge^3(\fg)^\fg .
\]
\end{cor}

In particular, when $\fg$ is semisimple the middle two terms in the above corollary vanish and we are left with a two-dimensional space of deformations $H^3_{\rm Lie}(\fg) \oplus \wedge^3 (\fg)^\fg = k \oplus k$. \brian{check me}

@@

This result is the simplest case of a generalized Hochschild-Kostant-Rosenberg theorem for commutative algebras viewed as $\EE_n$ algebras.
(The classic theorem proves the $n=1$ situation.)
Such a result is proved by very different methods in~\cite{CalWil, ToenBranes}.
Now we swiftly describe how to use methods like those in this paper to prove a version of this powerful result.

\owen{I've gotten myself pretty mixed up about this \dots too many duals with infinite dimensional vector spaces, and issues with whether you can base-change a semi-free algebra.}

First, we wish to define ${\rm Poly}_n(A)$ for an arbitrary dg commutative algebra $A$ in a way that gives the ``correct'' derived answer, i.e., the answer up to quasi-isomorphism of dg commutative algebras.
Concretely, one should replace $A$ by a quasi-isomorphic but better behaved algebra $\widetilde{A}$, 
such as semi-free resolution or cofibrant replacement with respect to some nice model category structure.
Then one computes the tangent complex $\TT_{\widetilde{A}}$, i.e., the derived derivations.
The {\em shifted polyvector fields} of $A$ means
\[
{\rm Poly}_n(A) \simeq \Sym_{\widetilde{A}} (\TT_{\widetilde{A}}[-n]),
\]
which is again a dg commutative algebra up to quasi-isomorphism.

%Second, we introduce a technical hypothesis  needed to apply our methods.
%We need an {\em augmented} dg commutative algebra $\epsilon: A \to k$ that admits a semifree resolution of the form $(\Sym(V^\vee), \partial)$, where $V$ is some graded vector space.
%In other words, we need that the generators can be realized as the graded linear dual of some other graded vector space.
%We say such an algebra $A$ is of {\em type~($\star$)}.

\begin{prp}
Let $A$ be an augmented dg commutative algebra.
Then 
\[
{\frak Z}_n(A) \simeq {\rm Poly}_n(A).
\]
\end{prp}

\owen{This is really turning into a mess.}

\begin{proof}
We are free to replace $A$ by a semifree resolution $(\Sym(V), \partial)$,
where $\partial$ sends $V$ to the augmentation ideal $\Sym^{> 0}(V)$.
Below we will (abusively) use $A$ to denote this semifree resolution,
as the original $A$ is irrelevant for us.
Ignoring the differential, we see that our situation resembles our central result for an abelian Lie algebra,
and so we now mimic the proof of our central result.

There is a natural homotopy-locally constant sheaf of cochain complexes $\Omega^\bu$ with differential $\d_{dR}$,
which is a soft replacement for the locally constant sheaf $\RR$.
Taking duals, we get a cosheaf.
In the same way, we define the factorization algebra 
\[
F_A = (\Sym((\Omega^\bu)^* \otimes V ), \d_{dR} + \partial) 
\]
which is manifestly commutative and is quasi-isomorphic to $A$ on every disk.
Hence $F_A$ is a commutative, locally constant factorization algebra that models~$A$.

Similar techniques allow us to construct a factorization algebra modeling the polyvector fields ${\rm Poly}_n(A)$.
Define
\[
Z_A = (\Hom^\bu(\Sym((\Omega^\bu)^* \otimes V), \Sym(\Omega_c^* \otimes V), [\d_{dR} + \partial,-]).
\]
Note its similarity to our model $\bZ_\fg$ for the center of~$\bU_n \fg$.
%There is a canonical shifted Poisson structure here by extending the canonical  pairing
%\[
%(\Omega^\bu \otimes V)^* \otimes \Omega^\bu_c \otimes V \to \RR
%\]
%arising from the inclusion $\Omega^\bu_c \otimes V \hookrightarrow \Omega^\bu \otimes V$.

Let us verify that 
\[
Z_A(D) \simeq {\rm Poly}_n(A)
\]
for any open disk $D$.
Due to the Poincar\'e lemma, we have canonical quasi-isomorphisms $(\Omega^\bu(D))^* \otimes V \xto{\simeq} V$ and $\Omega^\bu_c(D) \otimes V \xto{\simeq} V[-n]$,
where we view $V$ as simply a graded vector space but the de Rham complexes have the differential $\d_{dR}$.
(The differential $\partial$ will make its appearance shortly.)
Thus we have quasi-isomorphisms
\[
i: \Sym((\Omega^\bu(D))^* \otimes V) \xto{\simeq} \Sym(V)
\]
and 
\[
i_c: \Sym(\Omega_c^*(D) \otimes V) \xto{\simeq} \Sym(V[-n]).
\]
As $\partial$ acts only on the $V$-factors and not the de Rham factors,
we can modify all the complexes by adding $\partial$ to the differential.
Finally, if we fix a quasi-isomorphism 
\[
s: \Sym(V) \xto{\simeq} \Sym((\Omega^\bu(D))^* \otimes V),
\]
we get a map
\[
\begin{array}{ccc}
Z_A(D) & \to & \Hom^\bu(A,A)\\
\Phi & \mapsto & i_c \circ \Phi \circ s
\end{array}
\]
that is a quasi-isomorphism.
(Note that $s$ depends on the disk~$D$. 
We do not construct a map that intertwines with any structure maps of~$Z_A$.)

We now turn to constructing a map from this factorization algebra $Z_A$ to the center ${\frak Z}_n(A)$.
The key is to show that $Z_A$ fits into a commuting triangle as in Definition~\ref{D:centralizer},
because then we know $Z_A$ has a canonical map to~${\frak Z}_n(A)$.
This action is quite easy to see here: there is a canonical map
\[
\Hom^\bu(\Sym((\Omega^\bu)^* \otimes V), \Sym(\Omega_c^* \otimes V) \otimes \Sym((\Omega^\bu)^* \otimes V) \to \Sym((\Omega^\bu)^* \otimes V)
\]
that sends $\Phi \otimes x$ to $\iota(\Phi(x))$,
where 
\[
\iota: \Sym(\Omega_c^\bu \otimes V) \to \Sym((\Omega^\bu)^* \otimes V)
\]

4) Show this putative center fits into the commuting triangle and hence possesses a canonical map to the "true" center

5) Use the Francis-Lurie recognition principle to show this map is a quasi-iso
\end{proof}

@@

\begin{cor}
\label{cor: koszul dual to main theorem}
For any $L_\infty$ algebra $\fg$,
there is a quasi-isomorphism
\[
{\frak Z}_n(\clies(\fg)) \simeq \clies(\fg, \UU_n \fg).
\]
\end{cor}

@@

Now, when $n$ is odd, we apply the Chevalley restriction theorem to obtain an isomorphism of graded vector spaces
\[
\Sym(\fg[1-n])^G \cong \Sym \left(\fh[1-n]\right)^W .
\]

Similarly, when $n$ is even, we apply the discussion in the paragraph preceding the statement of this proposition 
\end{proof}

>>>>

The construction $\int_{S^{n-1}}\UU_n\fg$ produces a factorization algebra on $\RR$.
Indeed,
\begin{align*}
\int_{S^{n-1}}\UU_n\fg 
&\simeq \cliels(\Omega^\bu_c(\RR_{> 0} \times S^{n-1}) \otimes \fg)\\
&\simeq \cliels(\Omega^\bu_c(\RR_{> 0}) \otimes H^\bu(S^{n-1}) \otimes \fg)
\end{align*}
because $(n-1)$-sphere is formal over the rationals (or reals). 
We thus see that 
\[
\int_{S^{n-1}}\UU_n\fg \simeq \UU_1 (H^\bu(S^{n-1}) \otimes \fg)
\]
The cohomology of the $(n-1)$-sphere is 
\[
R_n = k[\epsilon_{n-1}]/(\epsilon_{n-1}^2),
\] 
where $|\epsilon_{n-1}| = n-1$.
Since $\UU_1 (H^\bu(S^{n-1}) \otimes \fg)$ is manifestly equivalent to the dg algebra~$U(\fg \otimes R_n)$,
we find that 
\[
\int_{S^{n-1}}\UU_n\fg \simeq U(\fg \otimes R_n)
\]
as $E_1$ algebras.
Recall as well that $\UU_n \fg$ is quasi-isomorphic to $\Sym(\fg[1-n])$ as cochain complexes.

The equivalence
\[
{\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n\fg}}(\UU_n \fg, (\UU_n \fh)_f) 
\simeq 
{\rm Mor}_{\LMod_{U(\fg \otimes R_n)}} (\Sym(\fg[1-n]), (\Sym(\fh[1-n]))_f)
\]
thus reduces the situation to familiar dg algebra.
By the inclusion of the unit $k \hookrightarrow R_n = k \oplus k[1-n]$,
we obtain a canonical inclusion $i: U\fg \to U(\fg \otimes R_n)$ of dg associative algebras,
and so
\[
k \otimes_{U\fg} U(\fg \otimes R_n) \simeq \Sym(\fg[1-n]).
\]
Base change along $i$ thus gives rise to an equivalence 
\[
{\rm Mor}_{\LMod_{U\fg}} (k, (\Sym(\fh[1-n]))_f) \simeq {\rm Mor}_{\LMod_{U(\fg \otimes R_n)}} (\Sym(\fg[1-n]), (\Sym(\fh[1-n]))_f)
\]
of underlying cochain complexes, i.e. an equivalence after applying the forgetful functor. 
The left hand side is modeled by $\clies(\fg,(\Sym(\fh[1-n]))_f)$, 
the Lie algebra cochains of $\fg$ with coefficients in~$(\UU_n \fh)_f$, as desired.
Hence the map $\theta_{\bf f}$ is a quasi-isomorphism on any disk, as needed.
\end{proof}

@@

When $n=1$, the above proposition recovers the well-known quasi-isomorphism
\[
{\rm Hoch}^\bu(U \fg , U \fg) \simeq \clie^\bu(\fg , \Sym(\fg)).
\]
See~\cite{LodayCyclicBook} for discussion of Hochschild (co)homology for such ``almost commutative'' algebras.

%As a corollary of the above proposition, we obtain a concrete description for $\EE_n$-deformations of $\UU_n \fg$ when $\fg$ is an ordinary (strict) Lie algebra. 
Consider now the situation of an ordinary Lie algebra~$\fg$.
When $\fg$ is semi-simple, the enveloping $\EE_1$ algebra $\UU_1 \fg$ is rigid in the sense that it admits no nontrivial deformations. 
This claim follows from the above equivalence (\ref{eqn:def}) together with the fact that $H_{\rm Lie}^2(\fg , \Sym(\fg)) = 0$ for $\fg$ is semi-simple by Whitehead's theorem.

For higher $n$, the $\EE_n$-enveloping algebra is no longer rigid, even in the case of semi-simple Lie algebras. 
For instance, consider the case $n=2$.

\begin{cor}
First-order deformations of $\UU_2 \fg$ are in bijective correspondence with
\[
H^3_{\rm Lie}(\fg) \oplus H^2_{\rm Lie}(\fg , \fg) \oplus H^1_{\rm Lie}(\fg , \wedge^2 \fg) \oplus \wedge^3(\fg)^\fg .
\]
In particular, for $\fg$ simple, the middle two terms vanish and deformations are isomorphic to $k \oplus k$.
\end{cor}

@@

This result is the simplest case of a generalized Hochschild-Kostant-Rosenberg theorem for commutative algebras viewed as $\EE_n$ algebras.
(The classic theorem proves the $n=1$ situation.)
Such a result is proved by very different methods in~\cite{CalWil, ToenBranes}.
Our method offers another approach, pursued in~\cite{GGWZ2}.

\subsection{Koszul duality}

@@

It will be convenient for us to prove a special case of that HKR theorem.

\begin{prp}
\label{cor: koszul dual to main theorem}
For any $L_\infty$ algebra $\fg$,
there is a quasi-isomorphism
\[
{\frak Z}_n(\clies(\fg)) \simeq {\frak Z}_n(\UU_n \fg) \simeq \clies(\fg, \UU_n \fg).
\]
\end{prp}

There is a structural feature to observe before we give the proof.
Note that $\clies(\fg)$ itself arises as a centralizer:
take the Lie algebra map $z: \fg \to 0$, 
in which case ${\frak Z}_n(z) \simeq \clies(\fg)$ because $\UU_n(0) \simeq 1$, the unit $\EE_n$ algebra.
The functoriality of the centralizer construction then tells us that ${\frak Z}_n(\UU_n \fg)$ acts on ${\frak Z}_n(\clies(\fg))$, by considering how the composite Lie algebra map 
\[
\fg \xto{\id} \fg \xto{z} 0
\]
determines an $\EE_n$ algebra map 
\[
\UU_n \fg \xto{\id} \UU_n \fg \xto{z} \UU_n(0) = 1
\]
and hence a map of centralizers
\[
{\frak Z}_n(\id) \otimes {\frak Z}_n(z) \to {\frak Z}_n(z).
\]
But ${\frak Z}_n(\id) = {\frak Z}_n(\fg)$ and ${\frak Z}_n(z) = \cliels(\fg)$.
In this way, we obtain a canonical map ${\frak Z}_n(\fg) \to {{\frak Z}_n(\clies(\fg))}$.
We wish to show it is an equivalence.

\begin{proof}
We use the factorization algebra $\clies(\Omega^* \otimes \fg)$ as is a factorization model for $\clies(\fg)$.
The idea is 
\begin{enumerate}
\item to show that our model $\bZ_n(\fg)$ for the center of $\bU_n \fg$ acts on the factorization algebra $\clies(\Omega^* \otimes \fg)$, and hence we obtain a canonical map $\bZ_n(\fg) \to {\frak Z}_n(\clies(\fg))$; and then
\item to use the Francis-Lurie recognition principle to verify this map is a quasi-isomorphism.
\end{enumerate}
Just before the proof, we explained why this canonical map exists for structural reasons,
and Theorem~\ref{T:compcentralizer} provides a completely explicit construction of it.

It remains to apply the recognition principle.
We compute
\begin{align*}
\int_{S^{n-1}}\clies(\fg) 
&\simeq \clies(\Omega^\bu(\RR_{> 0} \times S^{n-1}) \otimes \fg)\\
&\simeq \clies(H^\bu(S^{n-1}) \otimes \fg)
\end{align*}
because $(n-1)$-sphere is formal over the rationals (or reals). 
This pushforward to the real line $\RR$ is a locally constant factorization algebra,
which happens to be commutative here.
As the cohomology of the $(n-1)$-sphere is 
\[
R_n = k[\epsilon_{n-1}]/(\epsilon_{n-1}^2),
\] 
where $|\epsilon_{n-1}| = n-1$,
we find that 
\[
\int_{S^{n-1}}\clies(\fg) \simeq \clies(\fg \otimes R_n)
\]
as commutative algebras.

Consider now the equivalence
\[
{\rm Mor}_{\LMod_{\int_{S^{n-1}} \clies(\fg)}}(\clies(\fg), \clies(\fg)) 
\simeq 
{\rm Mor}_{\LMod_{\clies(\fg \otimes R_n)}} (\clies(\fg), \clies(\fg)).
\]
\owen{I think it's straightforward algebra after this (along the lines of the base change for the other proof), but I'm running out of steam right now.}
%The augmentation $R_n = k \oplus k[1-n] \to k$,
%induces a canonical inclusion $i: \clies(\fg) \to \clies(\fg \otimes R_n)$ of dg associative algebras,
%and so
%\[
%k \otimes_{U\fg} U(\fg \otimes R_n) \simeq \Sym(\fg[1-n]).
%\]
%Base change along $i$ thus gives rise to an equivalence 
%\[
%{\rm Mor}_{\LMod_{U\fg}} (k, (\Sym(\fh[1-n]))_f) \simeq {\rm Mor}_{\LMod_{U(\fg \otimes R_n)}} (\Sym(\fg[1-n]), (\Sym(\fh[1-n]))_f)
%\]
%of underlying cochain complexes, i.e. an equivalence after applying the forgetful functor. 
%The left hand side is modeled by $\clies(\fg,(\Sym(\fh[1-n]))_f)$, 
%the Lie algebra cochains of $\fg$ with coefficients in~$(\UU_n \fh)_f$, as desired.
%Hence the map $\theta_{\bf f}$ is a quasi-isomorphism on any disk, as needed.
\end{proof}

\owen{Add a discussion of Koszul duality. In the abelian case this is really obvious.}

@@

Now, when $n$ is odd, we apply the Chevalley restriction theorem to obtain an isomorphism 
\[
\Sym(\fg[1-n])^G \cong \Sym \left(\fh[1-n]\right)^W 
\]
of graded vector spaces.

Similarly, when $n$ is even, we apply the discussion in the paragraph preceding the statement of this proposition.
\end{proof}
